---
title: "D2SC ICMA Notebook"
output: html_notebook
author: "Vincent Espana"
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
```


# ICMA 9/9/2024 

```{r}
plot(cars)
```

# ICMA 9/11/2024

This is where we put plain text

```{r}
# this is where we put r code
```

```{r}
0hno!This_is.a_nightmare <- "hello"
```

```{r}
ohno_this_is_a_nightmare <- "hello"
# dots also work
```

```{r}
2+2
```
```{r}
x <- 2 + 2
```

```{r}
?mean
```

```{r}
mean(c(2, 3, 4))
```

```{r}
myVector <- c(2, 3, 4)
```

```{r}
mean(my_vector)
```
```{r}
my_vector <- c(2, 3, 4)
```

```{r}
?sd
```

```{r}
sd(1:2) ^ 2
```
# ICMA 9/16/2024

```{r}
#install.packages() last time
library(tidyverse)
```

```{r}
#some functions cant be used/need more work
stats::filter()
```

```{r}
#does not work
head(3, billboard)
```


```{r}
#whatever output is first is put into the first position
# 3 will technically be in second position
billboard %>%
  head(3)
```

```{r}
my_numbers <- c(1, 2, 3, 4, 5, 8, 9, 10)
```

```{r}
mean(my_numbers)
```
```{r}
my_numbers %>%
  mean()
```
```{r}
getwd()
```

```{r}
mtcars
```

```{r}
write_csv(mtcars, #object name
          "mtcars_fromR.csv") #file name to save

```

```{r}
mtcars %>%
  rownames_to_column() %>%
  write_csv("mtcars_fromR_rownames.csv")
```

```{r}
mtcars_read <- read_csv("mtcars_fromR_rownames.csv")
```
# ICMA 9/18/2024

```{r}
?ChickWeight
```

```{r}
ChickWeight %>%
  select(weight)
```

```{r}
chick_clean <- ChickWeight %>%
  select(chick_id = Chick,
         diet = Diet,
         age_days = Time,
         weight_gm = weight) #will reorder as well as rename
#will not print because this is only assignment --> will show up in environment
```

equality operator in R is "=="

```{r}
chick_clean %>%
  arrange(chick_id)
```

```{r}
chick_clean %>%
  mutate(weight_per_day = weight_gm/age_days)
```

```{r}
chick_clean %>%
  mutate(weight_lbs = weight_gm/453.6) #mutate will always add to end of dataset 
# doubles will go to infinite within computer bounds

```

```{r}
42/453.6
```
```{r}
chick_clean %>%
  mutate(weight_lbs = weight_gm/453.6) %>%
  group_by(diet, age_days) %>%
  summarize(n = n(),
            mean_wgt_gm = mean(weight_gm,),
            mean_wgt_lbs =mean(weight_lbs)) %>%
  filter(age_days == 20)
```

3 things:
for each diet group
number of chicks, avg gm weight, avg lbs weight
on day 20

```{r}
chick_clean %>%
  filter(age_days == 20) %>%
  mutate(weight_lbs = weight_gm/453.6) %>%
  group_by(diet) %>%
  summarize(n = n(),
            mean_wgt_gm = mean(weight_gm,),
            mean_wgt_lbs = mean(weight_lbs))
```

```{r}
mtcars_modified <- read_csv("mtcars_fromR_rownames.csv")
```

```{r}
glimpse(mtcars_modified)
```
# ICMA 9/23/2024

```{r}
billboard
```

```{r}
billboard %>% 
  pivot_longer(wk1:wk76)
```

```{r}
billboard_longer <- billboard %>%
  pivot_longer(cols = starts_with("wk"), 
               names_to = "week",
               values_to = "position",
               names_prefix = "wk",
               values_drop_na = TRUE)

billboard_longer
  
```

```{r}
billboard_longer %>% 
  pivot_wider(names_from = "week", 
              values_from = "position")
```

```{r}
?pivot_longer
```

```{r}
billboard %>% 
  separate(col = date.entered,
           into = c("year", "month", "day"))
```

```{r}
billboard %>%
  unite(col = "date",
        into = c(year, month, day))
```

# ICMA 9/25/2024

```{r}
ChickWeight %>%
  ggplot(aes(x = weight)) + # provide the aesthetic mapping 
   geom_histogram()
```

cant use pipe functions for geom histogram

```{r}
ChickWeight %>%
  ggplot(aes(y = weight,
             x = Time,
             color = Diet)) +
  geom_point() + geom_smooth()
```
kind of diet missing, need more information ^will add above
noise = random variability

```{r}

ChickWeight %>%
  ggplot(aes(y = weight,
             x = Time,
             color = Diet)) +
  geom_boxplot(alpha = 0.5) + # make a lil transparent
  #geom_smooth(se = FALSE) + #remove error bands
  theme_minimal() + # using a premade theme to change appearance
  facet_grid(~Diet) +
  labs(y = "weight in grams", 
       x = "days since birth",
       title = "Weight Over Time Across Diets")  

ggsave("chick_weightovertime.png", width = 9 , length = 8)
```

# ICMA 10/7/2024

types: numeric, character, logical 
shared functions: ypeof() - find type of object 
numeric: integers, doubles (r default)

doubles do not need decimals
be careful if you do as.integer("hello") -> will not give error will give warning, it will convert to NA (can't parse as number)

```{r}

as.numeric("12") #digits can parse into numbers

```

```{r}

my_double <- 12.9
typeof(my_double)
as.integer(my_double) #does not round, cuts off decimal part (truncate)
as.character(my_double)

```

NA introduced by coercion
NA is logical used as missing value indicator 
0 is a value, null is a null object

"missing" in sense of does not exist in R

```{r}

2==NA #result is NA (does not know) - does not know if it is equal to 2 because it does not know if it is 
NA==NA # us is.na()

my_values <- c(1,3,5,NA,8,10)
mean(my_values)
#typically can override this by adding a na.rm = TRUE argument to remove the NAs
mean(my_values, na.rm = TRUE) #ignore NA

```

single or/and will evaulate all
double or/and will evaulate until one 

prediction:
true
true
true 
true
true

```{r}

(3 / 4) > 0 && (8 + 1 - 7) == 2 # evaluates until one
(5 / 2) == 1 | (2 / 1) > 0 # evaluates all
(5 * (5-3)) == 10 || (10 / 2) < 4 # evaluates until one

is.numeric(mean(c(38,12,123.7,4.32)))
#TRUE would mean that the mean of those numbers is a numeric in R

((7/2) == 3.5) & TRUE | NA

```

```{r}

NA & TRUE # NA since NA is neither TRUE or FALSE
NA & NA # NA

```


