---
title: "D2SC ICMA Notebook"
author: "Vincent Espana"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
  word_document:
    toc: true
---

```{r}

knitr::opts_chunk$set(echo = FALSE)

```


# Initial Loading

```{r}

library(tidyverse)

```


# ICMA 9/9/2024 

```{r}
plot(cars)
```

# ICMA 9/11/2024

This is where we put plain text

```{r}
# this is where we put r code
```

```{r}
ohno_this_is_a_nightmare <- "hello"
# dots also work
```

```{r}
2+2
```
```{r}
x <- 2 + 2
```

```{r}
?mean
```

```{r}
mean(c(2, 3, 4))
```

```{r}
myVector <- c(2, 3, 4)
```

```{r}
my_vector <- c(2, 3, 4)
```

```{r}
?sd
```

```{r}
sd(1:2) ^ 2
```
# ICMA 9/16/2024

```{r}
#install.packages() last time
library(tidyverse)
```

```{r eval = FALSE}
#some functions cant be used/need more work
stats::filter()
```

```{r eval = FALSE}
#does not work
head(3, billboard)
```


```{r}
#whatever output is first is put into the first position
# 3 will technically be in second position
billboard %>%
  head(3)
```

```{r}
my_numbers <- c(1, 2, 3, 4, 5, 8, 9, 10)
```

```{r}
mean(my_numbers)
```
```{r}
my_numbers %>%
  mean()
```
```{r}
getwd()
```

```{r}
mtcars
```

```{r}
write_csv(mtcars, #object name
          "mtcars_fromR.csv") #file name to save

```

```{r}
mtcars %>%
  rownames_to_column() %>%
  write_csv("mtcars_fromR_rownames.csv")
```

```{r}
mtcars_read <- read_csv("mtcars_fromR_rownames.csv")
```
# ICMA 9/18/2024

```{r}
?ChickWeight
```

```{r}
ChickWeight %>%
  select(weight)
```

```{r}
chick_clean <- ChickWeight %>%
  select(chick_id = Chick,
         diet = Diet,
         age_days = Time,
         weight_gm = weight) #will reorder as well as rename
#will not print because this is only assignment --> will show up in environment
```

equality operator in R is "=="

```{r}
chick_clean %>%
  arrange(chick_id)
```

```{r}
chick_clean %>%
  mutate(weight_per_day = weight_gm/age_days)
```

```{r}
chick_clean %>%
  mutate(weight_lbs = weight_gm/453.6) #mutate will always add to end of dataset 
# doubles will go to infinite within computer bounds

```

```{r}
42/453.6
```
```{r}
chick_clean %>%
  mutate(weight_lbs = weight_gm/453.6) %>%
  group_by(diet, age_days) %>%
  summarize(n = n(),
            mean_wgt_gm = mean(weight_gm,),
            mean_wgt_lbs =mean(weight_lbs)) %>%
  filter(age_days == 20)
```

3 things:
for each diet group
number of chicks, avg gm weight, avg lbs weight
on day 20

```{r}
chick_clean %>%
  filter(age_days == 20) %>%
  mutate(weight_lbs = weight_gm/453.6) %>%
  group_by(diet) %>%
  summarize(n = n(),
            mean_wgt_gm = mean(weight_gm,),
            mean_wgt_lbs = mean(weight_lbs))
```

```{r}
mtcars_modified <- read_csv("mtcars_fromR_rownames.csv")
```

```{r}
glimpse(mtcars_modified)
```
# ICMA 9/23/2024

```{r}
billboard
```

```{r}
billboard %>% 
  pivot_longer(wk1:wk76)
```

```{r}
billboard_longer <- billboard %>%
  pivot_longer(cols = starts_with("wk"), 
               names_to = "week",
               values_to = "position",
               names_prefix = "wk",
               values_drop_na = TRUE)

billboard_longer
  
```

```{r}
billboard_longer %>% 
  pivot_wider(names_from = "week", 
              values_from = "position")
```

```{r}
?pivot_longer
```

```{r}
billboard %>% 
  separate(col = date.entered,
           into = c("year", "month", "day"))
```

```{r eval = FALSE}
billboard %>%
  unite(col = "date",
        into = c(year, month, day))
```

# ICMA 9/25/2024

```{r}
ChickWeight %>%
  ggplot(aes(x = weight)) + # provide the aesthetic mapping 
   geom_histogram()
```

cant use pipe functions for geom histogram

```{r}
ChickWeight %>%
  ggplot(aes(y = weight,
             x = Time,
             color = Diet)) +
  geom_point() + geom_smooth()
```
kind of diet missing, need more information ^will add above
noise = random variability

```{r eval = FALSE}

ChickWeight %>%
  ggplot(aes(y = weight,
             x = Time,
             color = Diet)) +
  geom_boxplot(alpha = 0.5) + # make a lil transparent
  #geom_smooth(se = FALSE) + #remove error bands
  theme_minimal() + # using a premade theme to change appearance
  facet_grid(~Diet) +
  labs(y = "weight in grams", 
       x = "days since birth",
       title = "Weight Over Time Across Diets")  

ggsave("chick_weightovertime.png", width = 9 , length = 8)
```

# ICMA 10/7/2024

types: numeric, character, logical 
shared functions: ypeof() - find type of object 
numeric: integers, doubles (r default)

doubles do not need decimals
be careful if you do as.integer("hello") -> will not give error will give warning, it will convert to NA (can't parse as number)

```{r}

as.numeric("12") #digits can parse into numbers

```

```{r}

my_double <- 12.9
typeof(my_double)
as.integer(my_double) #does not round, cuts off decimal part (truncate)
as.character(my_double)

```

NA introduced by coercion
NA is logical used as missing value indicator 
0 is a value, null is a null object

"missing" in sense of does not exist in R

```{r}

2==NA #result is NA (does not know) - does not know if it is equal to 2 because it does not know if it is 
NA==NA # us is.na()

my_values <- c(1,3,5,NA,8,10)
mean(my_values)
#typically can override this by adding a na.rm = TRUE argument to remove the NAs
mean(my_values, na.rm = TRUE) #ignore NA

```

single or/and will evaulate all
double or/and will evaulate until one 

prediction:
true
true
true 
true
true

```{r}

(3 / 4) > 0 && (8 + 1 - 7) == 2 # evaluates until one
(5 / 2) == 1 | (2 / 1) > 0 # evaluates all
(5 * (5-3)) == 10 || (10 / 2) < 4 # evaluates until one

is.numeric(mean(c(38,12,123.7,4.32)))
#TRUE would mean that the mean of those numbers is a numeric in R

((7/2) == 3.5) & TRUE | NA

```

```{r}

NA & TRUE # NA since NA is neither TRUE or FALSE
NA & NA # NA

```

# ICMA 10/9/2024

scalars, vectors, lists, data frames, tibbles

scalar: single value, like a number or string
 can me numeric (int, dbl), char, string
 
vector: a series of elements that have the same type
- numbers can be stored as strings, how to know --> quotations

- because vectors have to have the same type, R will coerce things to be the same type

```{r}

c("1", 2, 3 ,4) #takes value as a string

c(2, 3, 4) #takes values as ints

```
string has higher precedence than ints during coercion

lists: objects that can store multiple things of different types, including additional vectors --> use list()

```{r}

my_list <- list("letters" = c("A", "L", "J", "H"),
                "numbers" = 4:8,
                "more_numbers" = 10:30) 

my_list

```
single brackets [] will return the list
double brackets [[]] will return the values within the list; can also use $

data frame: each element is of the same length and are bound together --> will get error if diff length

use data.frame

also use [], [[]], and $ to access elements

tibble:a modern reimagining of the data.frame
- lazy and surly

create a tibble:
1. using a list of vectors using tibble()
2. type out row wise using tribble()

```{r}

age_data <- tribble(
                    ~id, ~age,
                    1, 8, 
                    2, 10, 
                    3, 8,
                    5, 9
  
)

gender_data <- tribble(
                        ~id, ~gender,
                        1, "f",
                        2, "m",
                        3, "nb",
                        4, "m",
                        6, "f"
)

age_data
gender_data

```

```{r}

full_join(age_data, gender_data)

```

```{r eval = FALSE}

bind_cols(age_data, gender_data)

```

```{r}

bind_rows(age_data, gender_data)

```

# ICMA 10/16/2024

ITERATIONS AND LOOPS
- often need to do the same thing over and over again
- functions help solve this program by packaging together a bunch of code
- loops help run the same code over and over again through iterations

```{r}

mean(c(1,2,3))

```
```{r}

(1+2+3)/3

```


```{r}

vec <- c(1,2,3)

(vec[1]+vec[2]+vec[3])/length(vec)

```
for loop iterates over a vector and executes R code after every iteration
general syntax:

```{r eval = FALSE}

for(var in seq){
  expr
}

```

example

```{r}

for (i in 1:10) {
  
  i <- i^2
  print(i)
}


```

```{r}

fave_number <- tribble(~name, ~number,
                       "MC", 8,
                       "AP", 9,
                       "NG", 12,
                       "VE", 6,
                       "AL", 17,
                       "AR", 3,
                       "MC", 7,
                       "RK", 8,
                       "AC", 4,
                       "SC", 21,
                       "KK", 16,
                       "RK", 28,
                       "VK", 17,
                       "MH", 12
                        )

```

1. for loop to print fave number

```{r}

for (i in fave_number$number) {
  
  print(i)
  
}

```

fave_number$number is different than fave_number["number"]... why?

```{r}

typeof(fave_number$number) #type of the vector
typeof(fave_number["number"]) #type of the list

```

one square bracket means pepper is in packaging in jar
two square brackt means actual pepper is in jar

```{r}

length(fave_number)

```
2. output for each person --> __'s fave number is __

```{r}

for (i in seq_along(fave_number$name)) {
  
  print(str_c(fave_number$name[i], "'s favorite number is ", fave_number$number[i]))
  
}

```

```{r}
?seq_along
```

```{r}

mtcars

```

everything(), map(), across()

while(cond) --> while something is true do this thing

# ICMA 10/30/2024

Let's write a sentence that has **bold** and *italics*.

Part 2: create a variable, then write a sentence that uses that variable, bold, and italics.

```{r}

name <- "Vincent"

```

Hi, my name is `r name` and I really have when people use both **bold** and *italics* in the same sentence.

```{r chick-plot, fig.cap = "Chick weight across time"}

ChickWeight %>%
  ggplot(aes(y = weight,
             x = Time, 
             group = Chick)) +
  geom_line()
  
```

# ICMA 12/2/2024

```{r}

months <- c("Apr", "Dec", "Jan", "Feb", "Mar")

```

```{r}

sort(months)

```

```{r}

months_base <- factor(months) #doesn't change the order of the actual thing, but will use levels when plotting
#default is alphabetical

```

```{r}

months_base <- factor(months, 
                 levels = c("Jan", "Feb", "Mar", "Apr", "May", "June", "July", "Dec"))

months

```

```{r}

months_forcats <- fct(months, 
                      levels = c("Jan", "Feb", "Mar", "Apr", "May", "June", "July", "Dec"))

months_forcats

```

if a level doesn't exist in the data set (example: removing "Dec") will make it an NA when changing it to factor
fct will through an error, factor will not