---
title: "Data Exercise 1: Decision Making"
author: "Vincent Espana"
date: "`r Sys.Date()`"
output: html_notebook
---

This is the notebook for the Data Exercise 1, about decision making. We did a study together in class on Qualtrics, then downloaded the data to analyze.

# Loading Libraries and Data

```{r}

library(tidyverse)

```

```{r}

de1 <- read_csv("CogSci412_DE1_data.csv") %>%
  filter(Progress == 100 & Status == "IP Address" ) %>%
  
  mutate(condition = case_when(is.na(FL_39_DO) == FALSE ~ FL_39_DO, 
                               is.na(FL_43_DO) == FALSE ~ FL_43_DO)) %>%
  
  select(ResponseId, 
         condition, 
         familiarity = Q10,
         pencil_ctrl = `Pencil Control`, 
         pencil_ctrl_conf = `Conf. Pencil Control`, 
         mag_stnd = `Magazine Standard`, 
         mag_stnd_conf = `Conf. Mag. Stand.` , 
         mag_control = Magazine_Control, 
         mag_ctrl_conf = `Conf. Mag. Control`, 
         pencil_stnd = Pencil_Standard, 
         pencil_stnd_conf = `Conf. Pencil Stand.`) %>%
  
  filter(ResponseId != "R_5R9cvvcJ5vFulLe")
  
  
 

```
This is in WIDE form, we want to get it in LONG form. How do we want to structure it?

Each cell can be categorized with three piece of information. What are they?
- pencil vs. magazine
- control vs. standard
- answer vs. confidence

What we want in long form:

ResponseId: unique identifier for that dataset
Content: what was the question about: pencils or magazines
Type: was the question a control question or a standard question
DV("dependent variable"): answer, confidence

value: the actual observed number/response

How do we get the wide form de1 into this desired format?

```{r}

long_de1 <- de1 %>%
  pivot_longer(pencil_ctrl:pencil_stnd_conf,
               names_to = "old_column",
               values_to = "value")

```


```{r}

long_de1_clean <- long_de1 %>% 
  filter(is.na(value) == FALSE) %>% 
  # check is.na() and is.na() == TRUE when the thing is NA and is == FALSE when the thing is not NA
  # in this case, we're keeping everything where the is.na(value) is NOT NA
  # therefore removing the ones where is.na(value) IS NA
  
  separate(into = c("content", "type", "DV"), 
            col = old_column) %>%
  
  # create new column, replacing the old one
  # NA is replaced with "resp"
  
  mutate(DV = case_when(is.na(DV) == TRUE ~ "resp",
                        DV == "conf" ~ "conf")) # will completely replace that column

```

To check whether our is.na(value) filter was working we did the following:
  group_by(ResponseId) %>%
  count() %>% 
  filter(n!= 4)
  
To interpret people's responses: 

for the pencil/eraser problem, for the standard correct answer is 5 cents, but we'd expect lots of folks to say 10cents. For the control probkem, the correct answer is 10 cents.

for the magazine/banana problems, the correct answer for the standard is 45 cents but we'd expect folks to say 90 cents. the correct answer for control is 90 cents

```{r}

long_de1_clean %>%
  filter(DV == "resp") %>%
  distinct(value)


```

```{r}

de1_analysis <- long_de1_clean %>%
  mutate(value = parse_number(value)) %>%
  mutate(value = case_when(value < 1 ~ value*100,
                           value >= 1 ~ value)) %>%
  mutate(correct_answer = case_when(content == "pencil" & type == "stnd" ~5,
                                    content == "pencil" & type == "ctrl" ~10,
                                    content == "mag" & type == "stnd" ~45,
                                    content == "mag" & type == "cntl" ~90),
         
         # create a new column for that person's accuracy on that trial
         # to see how if_else works, when you have one condition and then EVERYTHING ELSE
         
         accuracy = if_else(DV == "conf" ~ NA,
                            value == correct_answer ~ TRUE, # they are correct
                            value != correct_answer ~ FALSE) # they are incorrect 
  )
```

```{r}
de1_analysis %>%
  mutate(correct_answer = case_when(content == "pencil" & type == "stnd" ~5,
                                    content == "pencil" & type == "ctrl" ~10,
                                    content == "mag" & type == "stnd" ~45,
                                    content == "mag" & type == "cntl" ~90),
         
         # create a new column for that person's accuracy on that trial
         # to see how if_else works, when you have one condition and then EVERYTHING ELSE
         
         accuracy = if_else(DV == "conf" ~ NA,
                            value == correct_answer ~ TRUE, # they are correct
                            value != correct_answer ~ FALSE) # they are incorrect 
  )

```

```{r}

de1_analysis %>%
  filter(DV == "resp") %>%
  group_by(type, accuracy) %>%
  summarize(n_responses = n())
  ggplot(aes(x = type, y = n_responses, fill = accuracy)) + geom_bar(stat = "identity")
```

```{r}

de1_analysis %>%
  filter(DV == "resp" & type == "stnd" & accuracy == FALSE) %>% 
  ggplot(aes(x = as.factor(value))) + geom_bar() +
  facet_grid(~content)

```

