---
title: "D2SC: Weekly Assignments"
author: Vincent Espana
date: "`r Sys.Date()`"
output: html_notebook
---

# Initial Loading

```{r}

?library

```

```{r}

library(tidyverse)

```

# Week 1 Assignment

To figure out how to load a library in RStudio, I used a quick Google search. I entered "how to load libraries in R" into the searched bar and opened the first two links, both of which explained to use the "library()" function. Running an additional code block using the help function certified the contents of those link. After adding it into the code block and running it, I was able to see that the library was successfully loaded into RStudio. I reset R and cleared the outputs and ran it once more to be sure.

<https://bookdown.org/nana/intror/install-and-load-packages.html> <https://campus.datacamp.com/courses/intermediate-r/chapter-3-functions?ex=17>

```{r}

?tidyverse

```

Description

The 'tidyverse' is a set of packages that work in harmony because they share common data representations and 'API' design. This package is designed to make it easy to install and load multiple 'tidyverse' packages in a single step. Learn more about the 'tidyverse' at <https://www.tidyverse.org>.

# Week 2 Assignment

```{r}

MFID_analogy_read <- read_csv("/Users/vince/OneDrive/Documents/GitHub/Data2SciComm/WeeklyAssignments/tidy_data/MFIndD_analogy.csv")

```

```{r}

MFID_analogy_read

```

1.  The column that contains the unique identifier for each participant is the "qualtrics_id" column.

2a. There are 792 rows and 6 columns in the dataset. I did not use a specific code chunk for this information because the number of rows and columns are provided when I initially read in the csv file for number 1.

```{r}

MFID_analogy_read %>%
  distinct(qualtrics_id) %>%
  count()

```

2b. Using the distinct function, I was able to filter out repeated qualtrics ID's for multiple trials of the same respondee. Therefore, there are 99 individuals in the dataset, as seen in the code chunk above.

```{r}

MFID_analogy_read %>%
  group_by(qualtrics_id) %>%
  summarize(n_trials = n())

```

2c. As seen in the code chunk above, each of the 99 participants complete 8 trials, which was found using the count function n(). Therefore, every participant has the same number of trials.

# Week 3 Assignment

```{r}

MFID_summary <- MFID_analogy_read %>%
  group_by(qualtrics_id) %>%
  summarize(relational_matches = sum(response_type == "Rel"))

MFID_summary #make it show up
  
```

```{r}

MFID_summary %>%
  ggplot(aes(x = relational_matches)) +
  geom_histogram(binwidth = 1, fill = "gray", color = "black") +
  labs(title = "Distribution of Relational Matches", 
       x = "Number of Relational Matches", 
       y = "Number of Participants")


```

2.  The histogram shows that participants generally tend to choose either relational matches or object matches across all trials. This is derived from the fact that an overwhelming majority of participants are at both extremes of the histogram.

```{r}

MFID_analogy_read %>%
  group_by(qualtrics_id) %>%
  select(-match_type_0, -match_type_1, -response) %>% #taking additional info out
    pivot_wider(names_from = trial_number,
                values_from = response_type,
                names_prefix = "trial ") %>%
      select("trial 1", "trial 2", "trial 3", "trial 4", "trial 5", "trial 6", "trial 7", "trial 8") #reordering to make it look better

```

```{r, eval = FALSE}

MFID_analogy_read %>%
  group_by(qualtrics_id) %>%
    pivot_wider(names_from = trial_number,
                values_from = response_type) %>% 
      select(-match_type_0, -match_type_1, -response) %>% #remove these columns
      select(1, 2, 3, 4, 5, 6, 7, 8)

```

Show My Work

This code chuck did work, just not in the way that I intended it to. After removing the unnecessary columns, I wanted to reorder the columns so that it would start at trial 1 and increment to trial 8. In the corrected code chunk, I added a prefix of "trials" so that I would be more specific. I then realized that I am still a little unfamiliar with the syntax for certain functions, thus forgetting to put the column names in quotations.

# Week 4 Assignment

```{r}

MFID_REI_read <- read_csv("/Users/vince/OneDrive/Documents/GitHub/Data2SciComm/WeeklyAssignments/tidy_data/MFIndD_REI.csv")

head(MFID_REI_read)

MFID_REI_read

```

1.  It is likely that the columns for "response" and "scored_response" have different column types because they contain different data. The column type for response is <chr> because it is likely that the response can be more than just a number, resulting in the whole column being treated as a non-numeric character. It is likely that the column type for response is double because the column only contains numeric values.

```{r}

MFID_REI_read_fixed <- MFID_REI_read %>%
  mutate(response = case_when(response == 2 ~ 2,
                              response == 3 ~ 3,
                              response == 4 ~ 4,
                              response == "Strongly Agree" ~ 5,
                              response == "Strongly Disagree" ~ 1))
MFID_REI_read_fixed

```

```{r}

MFID_REI_read_mut <- MFID_REI_read_fixed %>% 
  mutate(new_scored_response = case_when(is.na(rev_scoring) == TRUE ~ response ,
                                          rev_scoring == "neg" ~ 6 - response))

MFID_REI_read_mut

```

```{r}

MFID_REI_read_mut %>%
  mutate(resp_check = case_when(new_scored_response == scored_response ~ TRUE, 
                                new_scored_response != scored_response ~ FALSE)
         ) %>% 
  distinct(resp_check)

```

# Week 5 Assignment

```{r}

MFID_REI_read_mut %>% distinct(sub_type) 

REI_summary <- MFID_REI_read_mut %>% 
  group_by(qualtrics_id, sub_type) %>%
    summarise(total_score = sum(new_scored_response)) 

REI_summary 

```

```{r}

na <- REI_summary %>% 
  filter(is.na(total_score))

na
  

```

1b. Yes, there are scores that are NA.

```{r}

REI_sum_notna <- MFID_REI_read_mut %>%
  group_by(qualtrics_id, sub_type) %>%
  summarise(total_score = sum(new_scored_response, na.rm = TRUE))

REI_sum_notna

na2 <- REI_sum_notna %>%
  filter(is.na(total_score))

na2

```

```{r}

MFID_combined <- REI_sum_notna %>%
  inner_join(MFID_summary)

MFID_combined

```

```{r}

ggplot(MFID_combined, aes(x = total_score, 
                          y = relational_matches,
                          color = sub_type)) + 
  geom_point() +
  geom_smooth() +
  labs(title = "Relation Between REI Score and Analogy Score", 
       x = "REI Total Score",
       y = "Analogy Score") + 
  theme_minimal()

```

3. There is some variability in the different best fit lines per sub-type. However,it seems that there is little to no correlation between the total scoring of the REI sub-types and the number of relational matches made. 

# Week 6 Assignment

1. loading data file

```{r}

MFID_prob <-MFID_analogy_read <- read_csv("/Users/vince/OneDrive/Documents/GitHub/Data2SciComm/WeeklyAssignments/tidy_data/MFIndD_probtask.csv")

MFID_prob

```

2a. create a vector

```{r}
conditions <- unique(MFID_prob$condition)

conditions # vector to store conditions

```
2b. calculate mean

```{r}

mean_rt <- numeric(length(conditions))

#filter invalid trials
valid_trials <- MFID_prob %>%
  filter(rt_exclu == 0)

for (i in 1:length(conditions)) {
  
  condition_data <- valid_trials[valid_trials$condition == conditions[i], ] #only keeping the columns that match the condition in conditions
  
  mean_rt[i] <- mean(condition_data$rt, na.rm = TRUE) #put the calculated mean in the mean_rt vector 
  
}

mean_rt

```
logical explanation: 

create a vector for the mean reaction times. in order to find the mean of each condition, i first need to filter out all of the valid trials (where rt_exclu == 0). then i can loop per condition and check where the condition in the valid trials matches one of the conditions in our conditions vector. after that i need to calculate the mean for all of the reactions of that specific condition and put that in our mean reaction time vector.

3. 

```{r}

summary_across <- valid_trials %>% 
  group_by(across(condition)) %>%
  summarize(mean_rt = mean(rt, na.rm = TRUE), # find the mean
            overall_accuracy = mean(correct, na.rm = TRUE)) # find the accuracy

summary_across

```

```{r}

summary_without <- summary_across <- valid_trials %>%
  group_by(condition) %>%
  summarize(mean_rt = mean(rt, na.rm = TRUE), # find the mean
            overall_accuracy = mean(correct, na.rm = TRUE)) #find the accuracy

summary_without

```

